### 2020-06-04 report/sections/server/node/challenges.tex
Report: complete chapter 4 sans connection

\section{Challenges}
\label{section:nodechallenges}
\begin{itemize}
\item session type implementation on the node runtime
\item need to provide consistency under the single-threaded event loop model
\item provide idiomatic event-driven APIs
\item provide static guarantees where possible -- communication mismatch and channel linearity
\item typescript's type system is gradual, structural and non-linear -- session type api generation examples that provide static guarantees on channel usage build upon target language type system's support for linear resources
\end{itemize}

As outlined in \cref{section:typescript}, TypeScript's type system
is neither affine nor linear, so we need to work around the 
limitations of the language to enforce linear usage of channel resources.
Similarly, we need to 
the language does not provide
built-in constructs for enforcing linear channel usage.

>>>>

\section{Challenges}
\label{section:nodechallenges}

Our goal with \fancyname{NodeMPST} is to
generate session type implementations that
provide as much \textbf{static} communication safety guarantees
as possible,
in a manner suitable for both the TypeScript language 
and the single-threaded event loop model in Node.js
As outlined in \cref{section:typescript}, TypeScript's type system
is neither affine nor linear, so we need to work around the 
language's limitations to enforce linear channel usage.

We also aim to provide the developer with
idiomatic event-driven APIs commonly found in
server-side TypeScript development.

### 2020-06-04 report/sections/server/node/limitations.tex
Report: complete chapter 4 sans connection

\begin{itemize}
\item verbose
\item state identifiers are in the API
\item scoping achieved through nested objects
\item performance aspect with class-based handlers
\end{itemize}

\begin{lstlisting}[language=javascript]
const first = new S4({ NUM1: (x) => secondOperand, });
const second = new S6({ NUM2: (y) => sum, });
const sum = [Labels.S7.SUM, ???, new S5()];
\end{lstlisting}

>>>>

Providing \textbf{static} communication safety guarantees have
come at a cost of providing relatively verbose APIs to the developer,
which introduces a learning curve. In particular,
the distinction between the \texttt{Implementation} and
\texttt{Handler} API should be an internal detail, but
the developer still needs to use the \texttt{Implementation} API
in their logic.

Our APIs also rely on state identifiers, which also
does not help code readability, as it is neither apparent nor
self-documenting what \texttt{S51} refers to -- some existing work
\cite{Hybrid2016} augment the state identifiers with the channel
action and labels involved (e.g. 
\texttt{S51_Receive__Add_Number__Quit_String}).

Using handler-style APIs also mean that we require
nested scoping to propagate values between states,
which does not scale elegantly as the levels of nesting
increase. For example, the following implementation of
the \tprotocol{One Adder} protocol would be invalid if
we define the handlers for the continuations separately,
because the received payload are no longer in scope
(\cref{line:notinscope}).

\begin{lstlisting}[language=javascript]
const first = new S4({ NUM1: (x) => secondOperand, });
const second = new S6({ NUM2: (y) => sum, });
const sum = [Labels.S7.SUM, (*@\label{line:notinscope}\hl{???}@*), new S5()];
\end{lstlisting}

However, the developer may choose to propagate values
between states through persistent storage APIs (such as
storing values in a database and retrieving them for
handlers of continuation states), so our generated APIs
do not strictly enforce scoping to be the only way to
propagate values in the application logic.

### 2020-06-04 report/sections/server/node/runtime.tex
Report: complete chapter 4 sans connection

	prepareReceive(
		next: EfsmTransitionHandler,
		register: (from: Roles.Peers, messageHandler: MessageHandler)
			=> void
	) {
		const messageHandler = (message: any) => {
			const decoded = JSON.parse(message) as Message.S51;
			switch (decoded.label) {
				case Labels.S51.ADD: {
					const successor = 
						this.handler[decoded.label](...decoded.payload);
					return next(successor);
				}
				case Labels.S51.QUIT: {
					// Same as above...		
				}
			}            
		}
		register(Roles.Peers.Client, messageHandler);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Generated Code for \texttt{Implementation} API for
Receive State}
\label{lst:nodereceive}
\end{figure}

>>>>

	prepareReceive(
		next: EfsmTransitionHandler,
		register: (from: Roles.Peers, messageHandler: MessageHandler)
			=> void
	) {
		const messageHandler = (message: any) => {
			const decoded = JSON.parse(message) as Message.S51;
			switch (decoded.label) {
				case Labels.S51.ADD: {
					const successor = 
						this.handler[decoded.label](...decoded.payload);
					return next(successor);
				}
				case Labels.S51.QUIT: {
					const successor = 
						this.handler[decoded.label](...decoded.payload);
					return next(successor);	
				}
			}            
		}
		register(Roles.Peers.Client, messageHandler);
	}
}
\end{lstlisting}
\captionof{lstlisting}{Generated Code for \texttt{Implementation} API for
Receive State}
\label{lst:nodereceive}
\end{figure}
